On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2015-09-02 MS.tex
Substantial edits throughout

\item
Traits should influence growth in a way that depend on individual plants size. We outline mechanisms and hypotheses based on new theory, and test these predictions in tree species for five traits using a meta-analysis of 112 studies (> 500 correlations). As a first approximation of plants size, we analysed the shift in the correlations between growth and traits across three plant stages (i.e. juvenile, sapling and adult).

@@

Currently, two main spectra of variation are widely cited as underpinning differences between species in vegetative growth rates. The ``leaf economic spectrum'' \citep{Wright:2004jb} and the ``wood economic spectrum'' \citep{Chave:2009iy} reflect tissue construction costs (for leaf area and for wood volume respectively), trading off with rates of tissue turnover or mortality. It has been widely expected that species with low tissue construction cost will typically have fast growth rates, at least in favourable physical environments \citep[e.g.][]{MullerLandau:2004dc,Wright:2004jb,Poorter:2008iu,Chave:2009iy,Larjavaara:2010bn,Iida:2012jb,Paine:2015df}. Yet, a growing number of empirical studies indicate the correlation between traits and growth is not as consistent as that. Consider SLA as an illustration. Species with high SLA can deploy more canopy area for a given dry mass investment in leaves \citep{Poorter:1999wd, Reich:1992wm}. In addition, SLA is correlated with high photosynthetic rates and high nitrogen (N) or phosphorus concentrations in the leaf \citep{Wright:2004jb}. For seedling relative growth rates under favourable conditions, high SLA has repeatedly been found to be a strong predictor regardless of vegetation type or growth form \citep{Lambers:1992bj,Reich:1992wm,Grime:1997wm,Poorter:1999wd,Wright:1999ds}. Yet for adult plants, this correlation appears weaker or absent \citep{coomes_comparison_1998,Poorter:2008iu,Aiba:2009ft,Easdale:2009gv,Wright:2010tp}. The growing discrepancy between theoretical expectation and empirical results has led some to question whether traits are even useful for understanding growth \citep{Wright:2010tp, Paine:2015df}.

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{Enquist:2007ek,Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}. Recent theory suggests that the absence of a correlation between SLA and growth at larger sizes may be expected, when potential influences of traits on plant function are considered \citep{Falster:2011ii, falster:2013}. At larger sizes the low-cost deployment benefits of high SLA may be offset or even outweighted by the high costs of rapid leaf turnover (see section ``Theoretical expectations''). While recent empirical studies supported the idea of a variation in the relationships between traits and growth rates across plant sizes (\citealt{Iida:2014ep, Iida:2014hq}), the trend and mechanisms behind this variation remain unclear. It is still an opening question whether the influence of traits on growth change with plant size in a predictable and consistent way.

In this paper, we used a meta-analytic approach on five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea, maximum height= Hmax, and seed mass= SM) to investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item  Do the observed relationships conform to predictions (outlined in section ``Theoretical expectations'') from a recently developed mechanistic model (Falster et al., 2011; Falster, 2013)?
\end{enumerate}
We gathered correlation coefficients between trait and growth from studies that gave explicit information about the size or stage of the plants. While most studies on traits have not been concerned with effects of plant size, they have generally reported trait-growth correlations for identifiable stage classes. We used plant stage (i.e. juvenile, sapling and adult) as a first approximation of plant size. We focused mainly on tree and woody species since the mechanistic model used to assess our predictions is most applicable to tree.

@@

Recent theoretical work \citep{Falster:2011ii} allows us to formulate hypotheses about relationships between traits and growth, based on a mathematical decomposition of plant growth rate. Specific predictions arise from considering the influence of traits on elements of these equations and how these effects may vary with size (see Appendix\ref{sec:growth} for mathematical explanations).

\textbf{H1: The relationship between specific leaf area (SLA) and growth rate is expected to shift from positive in seedlings to non-significant or negative in adult plants}

Plants with high SLA can deploy more leaf area per unit of mass invested. However, high-SLA species also have shorter leaf lifespan and faster leaf turnover rate \citep{Wright:2004jb}. Whether plant growth increases with SLA depends on the relative magnitude of the leaf construction and leaf turnover effects of SLA (Table 1). During early life, while leaf comprises a large part of the plant, decreasing leaf construction cost has a positive effect on growth rate. As plants increase in size and leaf mass fraction decreases, the benefit of cheaper leaf construction diminishes; thus the cost of replacing leaf turnover becomes more important.

\textbf{H2: The relationship between wood or stem density (WD) and growth rate is expected to be negative across all plant size}

Low WD decreases the cost of building stem tissue and thus the cost of deploying additional leaf area (Table 1), which should result in faster growth. Although cheaper wood carries a risk of increased mortality, this cost is not reflected in growth outcomes. Thus decreasing WD is predicted to increase growth across the entire size range.

\textbf{H3: The relationship between asymptotic maximum height (Hmax) and growth rate is expected to be absent in seedlings but to strengthen with increasing plant size.}

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do, Falster:2011ii, Wenk:2014jz}). For seedlings, where all individuals are investing solely in growth, we expect no influence of Hmax on growth. As plants become larger, species with smaller Hmax are expected to be allocating more to reproduction and less to growth, because at any given size they are closer to their Hmax (Table 1).

\textbf{H4: The relationship between seed mass (SM) and relative growth rate is negative in seedlings, but weakens with increasing size.}

The main effect of increased seed size is to produce a larger overall seedling. Generally relative growth rate decreases with size, and as such larger seedlings are predicted to have a lower relative growth rate during the earliest phases of growth \citet{Turnbull:2012ew}. However, among larger plants we expect the effect of this initial size to be minimal, and thus the correlation between seed size and growth rate should disappear.

\textbf{H5: The relationship between maximum photosynthetic rate (Aarea) and growth rate is expected to be positive across all plant sizes, at least under high-light conditions}

We expect greater photosynthetic capacity to translate into faster growth rates by increasing the rate of biomass production (Table 1), at least at high light where Aarea is measured. It is uncertain how much this might translate to faster growth in low light.

\textbf{Effect on absolute (AbsGR) and relative growth rate (RGR)}
Tree growth in diameter, height or mass can be expressed in either absolute (AbsGR) or relative terms (RGR). According to the growth model, the predicted relationship between growth and either SLA or seed mass in seedlings revert if absolute $Abs_{mass}$ is used instead of $RGR_{mass}$ (see Appendix 1\ref{sec:growth}). For SLA, the correlation with $RGR_{mass}$ is positive but becomes negative with $AbsGR_{mass}$. For SM the correlation with $RGR_{mass}$ is negative but becomes positive with $AbsGR_{mass}$. For the others traits (WD, Hmax and Aarea), the predicted relationships are not affected by the growth measurement used.

@@

\subsection*{Literature snapshot}\label{literature-snapshot}
We aimed to identify studies that measured the correlation between five prominent functional trait (SLA, WD, Aarea, Hmax and Seed mass) and a growth measurement across species. According to a snapshot of the literature on Web of Science database (Thomson Reuters), 82\% of the literature for these 5 traits is captured by searching only on SLA and WD. This snapshot was performed between March and June 2014 using a range of Boolean search terms in English. Search terms were as follows: (``TRAIT'' AND (``growth rate'' OR ``relative growth rate'')) AND (``tree*'' OR ``woody*''), ``TRAIT'' was all trait names listed in Table A1. While Web of Science guarantees a relatively stable and reproducible search environment, with clearly defined lists of indexed journals, its searches are restricted to metadata, such as title, keywords and abstracts \citep{Beckmann:2012hn}. This limitation does not apply to Google Scholar, where, theoretically, all text that is electronically available on the Internet will be indexed. Thus, we used Google Scholar to perform a more systematic search of the literature for SLA and WD.

\subsection*{Systematic search}\label{systematic-search}
Google Scholar database was searched using the range of Boolean search terms for SLA and WD. Search terms were as follows: ((``SLA'' OR ``LMA'' OR ``specific leaf area'' OR ``leaf mass area'' OR ``wood density'' OR ``WD'' OR ``wood specific gravity'' OR ``stem density'' OR ``stem specific density'') AND (``growth rate'' OR ``relative growth rate'')) AND (tree* OR woody*). The keyword search in Google Scholar returns approx. 18900 papers. For comparison, with the same keyword search Web of Science returns 541 papers. Thus, in Google Scholar we displayed papers by relevance and examined the first 750 items. We also identified and recorded other possible data sources using our own knowledge as researchers in the field (this included adding data from an as-yet unpublished study by Gleason et al). To avoid publication bias due to selective reporting of significant results within published articles (e.g.~particular case of a ``failure to publish bias'' see \citealt{Jennions:2013ta}), we asked the authors for additional information when they reported a non-significant correlation between growth and trait without reporting a correlation coefficient value. The complete list of studies where we inspected the full text is given in Appendix \ref{app:supp_info_files} in Supplementary material.

Each study had to meet some basic criteria to be included in our analysis. We restricted our meta-analysis to documents reporting: i) the relationship between growth and trait as a correlation coefficient, ii) the sample size and the number of species used to perform the correlation, and iii) the size/stage of plants. Ultimately data was included from 112 different articles (starred references in bibliography).

Where available we also recorded the geographic coordinates, number of sites, experiment type (greenhouse, field experiment, forest, shade-house etc.), growth measure (relative or absolute growth rate measured using plant diameter, height or mass), vegetation type, life form (tree, woody or mixed with herbaceous species etc.), biological scale (interspecific, intraspecific or intragenus), and experimental treatments (e.g. fertilization, shade, salt stress). We reported coefficient correlations for a list of 20 traits (Fig. \ref{Fig1}): the five for which we had prior expectations, plus 15 others which capture different aspects of plant species strategies and are commonly addressed in the trait literature. Specific leaf area (SLA), leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and assimilation rate of CO2 (Aarea) are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. Seed mass (SM), and maximum plant height (Hmax) capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. Wood density (WD) and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy  \citep{Chave:2009iy}. Here WD and SA/LA are also considered as indicators of stem construction cost (\citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource allocation to leaf \citep{Lambers:1992bj}. According to the snapshot of the literature on Web of Science, 62\% of the literature for these 20 traits were captured by a search on SLA and WD.

\subsection*{Plant stage categories}\label{plant-stage-categories}

The authors of the 112 articles reported in our meta-analyses, categorized plants into four differents stages (i.e. seedling, juvenile, sapling, adult). While the stage categories have been established according to the expertise and the appreciation of the authors, by compiling data across multiple studies and across different growth form (tree and woody, etc), there is a risk of assigning plants of similar size to different categories. We verified the consistency of size/age range attributed to each stage across the 112 studies reported here. Fig. \ref{Fig2} showed that the thresholds in size/age used by the authors to defined a plant stage are quite steady. In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m); there was also a clear break between saplings and adults in diameter (at 10 cm). According to simple allometric relationships from the Biomass And Allometry Database (BAAD) \citep{Falster:2015}, these stage cutoffs match across the different size measurements (i.e. diameter, height and age, Fig. \ref{FigA1}). The size/stage threshold defined across the 112 articles, allowed us to attribute straightforwardly a stage category to studies for which no stage have been distinctly mentionned by the authors (5\% of the data, noted as ``mix'' in Fig. \ref{Fig2}).

Less than 5\% of the studies (i.e. five studies on 112) were ambiguous on their plant stage attribution. Half of these equivocal cases involved the juvenile category. The term ``juvenile'' is vague enough to cover size range from either seedling or sapling stages. according to the authors. This redundancy encourages us to focus on a three stage classification (i.e. juvenile, sapling, adult) instead of a four stage one (i.e. seedling, juvenile, sapling, adult). In the rest of the mansucript we preferred the term ``juvenile'' to ``seedling'', because for a portion of the scientific community ``seedling'' also refers to a particular experimental condition (a greenhouse experiment). We attributed a study to a new stage category only when 95\% of the recorded range of sizes for that study lay within a new stage category. The others equivocal cases were caused by an ambiguity in the definition of ``adult'' used by the authors. Some plants with a large stature are not assessed as adult by the authors because they are still non reproductive (see ID 52 and 50). Here, we attributed a study to the adult stage when the authors identified the plants as reproductive, and to sapling stage when they identified the plants as non-reproductive, regardless the plant size or age. Results using the categories established by the authors are similar than the one using the standardized categories and are available in Fig. \ref{FigA5} model.1. In the main text, we report the results using the standardized categories.

@@

Effect size calculation:
We used the cross-species Pearson product-moment correlation coefficient r between growth rate and species trait as our measure of effect size, weighted by the number of species contributing. When studies did not report r values but another coefficient such as coefficient of determination (R2), Spearman's rho rank correlation or Kendall's tau rank correlation, we converted these to correlation coefficients  \citep{Lajeunesse:2013tm} (see code in Appendix \ref{app:supp_info_files} for details). When the magnitude of Pearson's correlation coefficient approaches +/- 1, its distribution becomes skewed. Accordingly we transformed the correlation r into a metric with desirable statistical properties using Fisher's z-transformation,

\[ z = \frac{1}{2} \ln \frac{1+r}{1-r} \]

which has a variance estimate

\[ v_z= \frac{1}{n-3} \]

n being the number of species. We weighted effect size by the study precision using the number of species instead of the sample size because the question addressed in our study concerned the correlation between growth and traits across species and not within species.

Stage effect:
We tested stage effect on traits for which > 15 correlation coefficients across at least two stages were recorded. We used a mixed-effect model of meta-analysis, in which it is assumed that studies within a stage share a common mean effect but that there is also random variation among studies in a stage, in addition to sampling variation. This random effect allows us to account for the heterogeneity among studies (e.g.~experimental type, vegetation type, error, etc.). For each trait, we established if stage could explain variance in the coefficient correlation by comparing models with the stage variable to the null model using log-likelihood ratios \citep{Zuur:2009cfa}. If the log-likelihood test returned a significant p-value, the stage model fitted the data better than the null model. Estimates of the effect size for each stage were considered significant if the 95\% confidence intervals did not overlap with zero.

Sensitivity analysis:
We tested how robust are our results to the decision made in the analysis. In particular we tested our prediction using two datatset with different inclusion criteria. The number of species used to perform the correlation, the experimental conditions, and the concordance between stage used to measure trait and growth are clear criteria for assessing the quality of the data for addressing our hypotheses. Therefore, we focused exclusively on correlation across species, and we stratified the data set into two subsets: one including only studies under unstressed conditions ('ideal' dataset), and the other set including studies under all conditions ('complete' dataset). We excluded from the ``ideal'' dataset studies where growth and traits were measured on plants from different stages, unless a particular stage is by definition required to measure a trait such as for seed mass (i.e. seed stage), or Hmax (i.e. adult stage).
We also tested the sentivity of our results to the dominance of one or a few studies that measured a large number of response for the same trait. In the case of multiple comparisons within a study, we used the method proposed by \citet{Borenstein:2009um}: we calculated the mean correlation coefficient by trait and by study, and re-calculated the effect size. This corresponds to the ”reduced dataset” where multiple responses were condensed into a single average within a trait and within a study.

Quality criteria for meta-analysis:
We conducted several analyses to assess the reliability of our meta-analysis. In particular, we quantified heterogeneity in effects size, we tested for publication bias and we explored temporal change in effect size (see Appendix \ref{app:supp_info_files}). We addressed the 14 methodological quality criteria listed by \citet{Koricheva:2014ku} as required to assess the relaibility of a meta-analyses (Table A2 detailed how and where in this article each citeria is addressed).

Availability of code and data:
All analyses were conducted with R software \citep{Ralanguageanden:2014wf}. The code and data underpinning this paper is available at github.com/AnaisGibert/Growth\_trait\_metaanalysis. Archival copies are also included in Appendix \ref{app:supp_info_files} in Supplementary material.

@@

For both SLA and SM, the correlation between growth rate and the trait changed significantly with plant size in the manner expected. Model fits improved when stage was incorporated (see log likelihood ratio tests, Fig. \ref{Fig3} a and d; average values for the coefficients r ($r_{\textrm{mean}}$) in Fig. \ref{FigA4}). The correlation between growth and SLA shifted from strongly positive at juvenile stage $r_{\textrm{mean}}=0.6 \pm 0.08$, Mean +SD) to not significantly different from zero at adult stage ($r_{\textrm{mean}}= -0.01 \pm 0.07$, CI 95\% overlapping zero in Fig. \ref{Fig3} a). The correlation between growth and SM shifted from negative at juvenile stage ($r_{\textrm{mean}}= -0.59 \pm 0.07$) to not significantly different from zero at adult stage ($r_{\textrm{mean}}= -0.09 \pm 0.02$, , CI 95\% overlapping zero in Fig. \ref{Fig3} d). For both traits, correlation coefficients recorded at sapling stage were intermediate between juvenile and adult stages.

>>>>

\item
Traits should influence growth in a way that depend on individual plants size. We outline mechanisms and hypotheses based on new theory, and test these predictions in tree species for five traits using a meta-analysis of 112 studies ($>$ 500 correlations). As a first approximation of plants size, we analysed the shift in the correlations between growth and traits across three plant stages (i.e. juvenile, sapling and adult).

@@

Currently, two main spectra of variation are widely cited as underpinning differences between species in vegetative growth rates. The ``leaf economic spectrum'' \citep{Wright:2004jb} and the ``wood economic spectrum'' \citep{Chave:2009iy} reflect tissue construction costs (for leaf area and for wood volume respectively), trading off with rates of tissue turnover or mortality. It has been widely expected that species with low tissue construction cost will typically have fast growth rates, at least in favourable physical environments \citep[e.g.][]{MullerLandau:2004dc,Wright:2004jb,Poorter:2008iu,Chave:2009iy,Larjavaara:2010bn,Iida:2012jb,Paine:2015df}. Yet, a growing number of empirical studies indicate the correlation between traits and growth is not as consistent as that. Consider SLA as an illustration. Species with high SLA can deploy more canopy area for a given dry mass investment in leaves \citep{Poorter:1999wd, Reich:1992wm}. In addition, SLA is correlated with high photosynthetic rates and high nitrogen (N) or phosphorus concentrations in the leaf \citep{Wright:2004jb}. For seedling relative growth rates under favourable conditions, high SLA has repeatedly been found to be a strong predictor regardless of vegetation type or growth form \citep{Lambers:1992bj,Reich:1992wm,Grime:1997wm,Poorter:1999wd,Wright:1999ds}. Yet for adult plants, the correlation is mostly much weaker or absent \citep{coomes_comparison_1998,Poorter:2008iu,Aiba:2009ft,Easdale:2009gv,Wright:2010tp}. The growing discrepancy between theoretical expectation and empirical results has led some to question whether traits are even useful for understanding growth \citep{Wright:2010tp, Paine:2015df}.

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}. Recent theory suggests that the absence of a correlation between SLA and growth at larger sizes may be expected, when potential influences of traits on plant function are considered \citep{Falster:2011ii}. At larger sizes the low-cost deployment benefits of high SLA may be offset or even outweighed by the high costs of rapid leaf turnover (see section ``Theoretical expectations''). While recent empirical studies supported the idea of a variation in the relationships between traits and growth rates across plant sizes (\citealt{Iida:2014ep, Iida:2014hq}), the trend and mechanisms behind this variation remain unclear. Moreover, the question whether the influence of traits on growth changes with plant size in a predictable and consistent way is very much open.

In this paper, we used a meta-analytic approach on five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea, asymptotic or maximum height = Hmax, and seed mass = SM) to investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item  Do the observed relationships conform to predictions (outlined in section ``Theoretical expectations'') derived from a recently developed mechanistic model \citep{Falster:2011ii}?
\end{enumerate}
We gathered correlation coefficients between trait and growth from studies that gave explicit information about the size or stage of the plants. While most studies on traits have not been concerned with effects of plant size, they have generally reported trait-growth correlations for identifiable stage classes. We used plant stage (i.e. juvenile, sapling and adult) as a first approximation of plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.

@@

Recent theoretical work \citep{Falster:2011ii} allows us to formulate hypotheses about relationships between traits and growth, based on a mathematical decomposition of plant growth rate. Specific predictions arise from considering the influence of traits on elements of these equations and how these effects may vary with size (see Appendix \ref{sec:growth} for mathematical explanations). This model extends earlier efforts to model growth as a function of traits \citep{Lambers:1992bj,cornelissen_seedling_1996,Wright:2000kw,Enquist:2007ek}. However, whereas earlier models focused solely on mass-based relative growth rate and did not predict a change in effect of traits with size, the revised equations do allow for these effects, while also making predictions for both absolute and relative growth rates in any of height, diameter or mass.

\textbf{H1: The relationship between SLA and growth rate is expected to shift from positive in seedlings to non-significant or negative in adult plants}

Plants with high SLA can deploy more leaf area per unit of mass invested. However, high-SLA species also have shorter leaf lifespan and faster leaf turnover rate \citep{Wright:2004jb}. Whether plant growth increases with SLA depends on the relative magnitude of the leaf construction and leaf turnover effects of SLA (Table \ref{tab:trade-offs}). During early life, when leaf comprises a large part of the plant, decreasing leaf construction cost has a positive effect on growth rate. As plants increase in size and leaf mass fraction decreases, the benefit of cheaper leaf construction diminishes; thus the cost of replacing leaf turnover becomes more important.

\textbf{H2: The relationship between WD and growth rate is expected to be negative across all plant size}

Low WD decreases the cost of building stem tissue and thus the cost of deploying additional leaf area (Table \ref{tab:trade-offs}), which should result in faster growth. Although cheaper wood carries a risk of increased mortality, this cost is not reflected in growth outcomes. Thus decreasing WD is predicted to increase growth across the entire size range.

\textbf{H3: The relationship between Hmax and growth rate is expected to be absent in seedlings but to strengthen with increasing plant size.}

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do,Thomas-2011, Wenk:2014jz}). For seedlings, where all individuals are investing solely in growth, we expect no influence of Hmax on growth. As plants become larger, species with smaller Hmax are expected to be allocating more to reproduction and less to growth, because at any given size they are closer to their Hmax (Table \ref{tab:trade-offs}).

\textbf{H4: The relationship between SM and relative growth rate is negative in seedlings, but weakens with increasing size.}

The main effect of increased seed size is to produce a larger overall seedling. Generally relative growth rate decreases with size, and as such larger seedlings are predicted to have a lower relative growth rate during the earliest phases of growth \citep{Turnbull:2012ew}. However, among larger plants we expect the effect of this initial size to be minimal, and thus the correlation between seed size and growth rate should disappear.

\textbf{H5: The relationship between Aarea and growth rate is expected to be positive across all plant sizes, at least under high-light conditions}

We expect greater photosynthetic capacity to translate into faster growth rates by increasing the rate of biomass production (Table \ref{tab:trade-offs}), at least at high light where Aarea is measured. It is uncertain how much this might translate to faster growth in low light.

\textbf{Effect on absolute and relative growth rates}
Tree growth in diameter, height or mass can be expressed in either absolute (AGR) or relative terms (RGR). According to the growth model, the predicted relationship between growth and either SLA or SM in seedlings revert if absolute AGR$_\textrm{mass}$ is used instead of RGR$_\textrm{mass}$ (see Appendix \ref{sec:growth}). For SLA, the correlation with RGR$_\textrm{mass}$ is positive but becomes negative with AGR$_\textrm{mass}$. For SM the correlation with RGR$_\textrm{mass}$ is negative but becomes positive with AGR$_\textrm{mass}$. For the others traits (WD, Hmax and Aarea), the predicted relationships are not affected by the growth measurement used.

@@

\subsection*{Dataset construction}

We aimed to identify studies that measured the correlation between five prominent functional trait (SLA, WD, Aarea, Hmax and SM) and a growth measurement across species. Our literature search was conducted in two phases.

\subsubsection*{Phase 1 - Literature snapshot}\label{literature-snapshot}

We first developed a snapshot of the literature, to assess potential to recover useufl data from the wide corpus of literature available. This snapshot was obtained via Web of Science database (Thomson Reuters), between March and June 2014, using a range of Boolean search terms in English. Search terms were as follows: (``trait'' AND (``growth rate'' OR ``relative growth rate'')) AND (``tree*'' OR ``woody*''), where ``trait'' was allowed to be any of the trait names listed in Table \ref{tab:lit_search}. This snapshot indicated that 82\% of the reported correlations linking growth to any of our five core traits would be recovered by searching only on SLA and WD, allowing us to narrow our search considerably.

\subsubsection*{Phase 2 - Systematic search}\label{systematic-search}

While Web of Science guarantees a relatively stable and reproducible search environment, with clearly defined lists of indexed journals, its searches are restricted to meta-data, such as title, keywords and abstracts \citep{Beckmann:2012hn}. This limitation does not apply to Google Scholar, where, theoretically, all text that is electronically available on the Internet will be indexed. Thus, we used Google Scholar to perform a more systematic search of the literature for SLA and WD.

The Google Scholar database was searched using a range of Boolean search terms for SLA and WD: ((``SLA'' OR ``LMA'' OR ``specific leaf area'' OR ``leaf mass area'' OR ``wood density'' OR ``WD'' OR ``wood specific gravity'' OR ``stem density'' OR ``stem specific density'') AND (``growth rate'' OR ``relative growth rate'')) AND (``tree*'' OR ``woody*''). This keyword search in Google Scholar returned approx. 18900 papers, from which we took the first 750 items when ranked by ``relevance''.
These papers were examined in addition to those discovered in Phase 1. In addition, we gleaned data from other sources, using our own knowledge as researchers in the field (this included adding data from an as-yet unpublished study by Gleason et al).  The complete list of studies where we inspected the full text is given in Appendix \ref{app:supp_info_files} in Supplementary material.

Each study had to meet some basic criteria to be included in our analysis. We restricted our meta-analysis to documents reporting: i) the relationship between growth and trait as a correlation coefficient, ii) the sample size and the number of species used to perform the correlation, and iii) the size/stage of plants. Ultimately data was included from 112 different articles (starred references in bibliography). Where available, we also recorded the geographic coordinates, number of sites, experiment type (greenhouse, field experiment, forest, shade-house etc.), growth measure (relative or absolute growth rate measured using plant diameter, height or mass), vegetation type, life form (tree, woody or mixed with herbaceous species etc.), biological scale (interspecific, intraspecific or intragenus), and experimental treatments (e.g. fertilization, shade, salt stress). To minimise publication bias via selective reporting of significant results within the published articles (e.g. a particular case of a ``failure to publish bias'' see \citealt{Jennions:2013ta}), we asked those authors reporting a non-significant correlation for additional information about the specific correlation coefficient when these were not provided.

Overall, we recorded coefficient correlations for a list of 20 traits (Fig. \ref{fig:fig1}): the five for which we had prior expectations, plus 15 others commonly used in the trait literature. SLA, leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and Aarea are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. SM and Hmax capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. WD and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy  \citep{Chave:2009iy}. Here WD and SA/LA are also considered as indicators of stem construction cost (\citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource allocation to leaf \citep{Lambers:1992bj}. According to the snapshot of the literature on Web of Science, 62\% of the literature for these 20 traits were captured by a search on SLA and WD.

\subsubsection*{Plant stage categories}\label{plant-stage-categories}

We were unable to obtain a consistent and continuous measure of plant size across all studies, thus we relied on categorical groupings by ``stage'' to test our hypotheses. The authors of the 112 articles used in the meta-analyses, reported their plants as belonging to any of fix different stages: ``seedling'', ``juvenile'', ``sapling'', ``adult'', ``mixed''. While these categories have been established according to the expertise of the authors, by compiling data across multiple studies and vegetation types, we were concerned about the risk of assigning plants of similar size to different categories. This risk turned out to be negligible: overall, we found that the original thresholds in size/age used by the authors to distinguish different stages was relatively compatible (Fig. \ref{fig:fig2}). In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m); there was also a clear break between saplings and adults in diameter (at 10 cm). According to simple allometric relationships between size variables plotted using data from the Biomass And Allometry Database \citep{Falster:2015}, these thresholds also align relatively well with one another (Fig. \ref{fig:figA1}). Thus we adopted the above values as cut-offs for defining stage boundaries. However, the groups ``juvenile'' and ``sapling'' were largely overlapping, causing us to merge these two.

Using the identified boundaries among stages we were able to assign all studies into one of three classes: ``juvenile'', ``sapling'', or ``adult''.  We preferred the term ``juvenile'' to ``seedling'', because for a portion of the scientific community ``seedling'' also refers to a particular experimental condition. Using the above cut-offs we were also able to assign a stage category to studies for which no stage have been distinctly mentioned by the authors (5\% of the data, noted as ``mix'' in Fig. \ref{fig:fig2}). Only 5 of 112 studies were ambiguous on their plant stage attribution. Half of these equivocal cases involved the juvenile category. We attributed a study to a new stage category only when 95\% of the recorded range of sizes for that study lay within a new stage category. The others equivocal cases were caused by an ambiguity in the definition of ``adult'' used by the authors. Some plants with a large stature are not assessed as adult by the authors because they are still non reproductive (see ID 52 and 50). Here, we attributed a study to the adult stage when the authors identified the plants as reproductive, and to sapling stage when they identified the plants as non-reproductive, regardless the plant size or age. Results using the categories established by the authors are similar than the one using the standardized categories and are available in Fig. \ref{fig:figA5} model.1. In the main text, we report only results using the standardized categories.

@@

\subsubsection*{Effect size calculation}

As a measure of effect size, we used the cross-species Pearson product-moment correlation coefficient $r$ between growth rate and trait, weighted by the number of species contributing to the relationship. For studies that reported another coefficient -- such as coefficient of determination, Spearman's rho rank correlation, or Kendall's tau rank correlation -- we converted these to correlation coefficients following \citet{Lajeunesse:2013tm}. As the magnitude of Pearson's correlation coefficient approaches $\pm$ 1.0, its distribution becomes skewed. Accordingly we transformed the correlation $r$ into a metric $z$ with more-desirable statistical properties. The transformation was achieved via Fisher's z-transformation,
\[ z = \frac{1}{2} \ln \frac{1+r}{1-r}, \]
which has a variance estimate of
\[ v_z= \frac{1}{n-3}, \]
where $n$ is the number of species. We weighted effect size by the number of species used to calculation the correlation, instead of the raw number of individuals, because our questions primarily concerned the correlation between growth and traits across species and not within species.

\subsubsection*{Stage effect}

We tested stage effect on traits for which $>$ 15 correlation coefficients across at least two stages were recorded. We used a mixed-effect model of meta-analysis, in which it is assumed that studies within a stage share a common mean effect but that there is also random variation among studies in a stage, in addition to sampling variation. This random effect allows us to account for the heterogeneity among studies (e.g. experimental type, vegetation type, error, etc.). For each trait, we established if stage could explain variance in $z$ by comparing models with the stage variable to the null model using log-likelihood ratios \citep{Zuur:2009cfa}. If the log-likelihood test returned a significant p-value, the stage model fitted the data better than the null model. Estimates of the effect size for each stage were considered significant if the 95\% confidence intervals did not overlap with zero.

\subsubsection*{Sensitivity analysis}

We tested how robust are our results were to the particular decisions made in the analysis, in the criteria used for including a study in the analysis. The number of species used to perform the correlation, the experimental conditions, and the concordance between stage used to measure trait and growth are clear criteria for assessing the quality of the data for addressing our hypotheses. Thus we stratified the data set into two subsets: one including only studies under unstressed conditions ('ideal' dataset), and the other set including studies under all conditions ('complete' dataset). We excluded from the ``ideal'' dataset studies where growth and traits were measured on plants from different stages, unless a particular stage is by definition required to measure a trait such as for SM (i.e. seed stage), or Hmax (i.e. adult stage).
We also tested the sensitivity of our results to inclusion of the few studies that measured a large number of responses for the same trait. In the case of multiple comparisons within a study, we used the method proposed by \citet{Borenstein:2009um}: we calculated the mean correlation coefficient by trait and by study, and then re-calculated the effect size. This produced a reduced dataset, where multiple responses for any given trait were condensed into a study-level average.

\subsubsection*{Quality criteria for meta-analysis}

\cite{Koricheva:2014ku} outlined 14 methodological quality criteria to ensure robust results from meta-analyses. Table \ref{tab:meta} outlines how we address each of these.  In particular, we quantified heterogeneity in effects size, we tested for publication bias and we explored temporal change in effect size (see Appendix \ref{app:supp_info_analyses}).

\subsection*{Availability of code and data}\label{code}

All analyses were conducted with R software \citep{Ralanguageanden:2014wf}. The code and data for producing all figures and results in this paper is available at github.com/AnaisGibert/Growth\_trait\_metaanalysis.

@@

For both SLA and SM, the correlation between growth rate and the trait changed significantly with plant size in the manner expected. Model fits improved when stage was incorporated (see log likelihood ratio tests, Fig. \ref{fig:fig3} a and d; average values for the coefficients ($\bar{r}$) in Fig. \ref{fig:figA4}). The correlation between growth and SLA shifted from strongly positive at juvenile stage $\bar{r}=0.6 \pm 0.08$, Mean $\pm$ SD) to not significantly different from zero at adult stage ($\bar{r}= -0.01 \pm 0.07$, CI 95\% overlapping zero in Fig. \ref{fig:fig3} a). The correlation between growth and SM shifted from negative at juvenile stage ($\bar{r}= -0.59 \pm 0.07$) to not significantly different from zero at adult stage ($\bar{r}= -0.09 \pm 0.02$, , CI 95\% overlapping zero in Fig. \ref{fig:fig3} d). For both traits, correlation coefficients recorded at sapling stage were intermediate between juvenile and adult stages.
